Comparing vector-based and ACT-R memory models of tag retrieval:\protect\\User-customized hashtag/tag prediction on Twitter and StackOverflow

### 2014-07-29 writeups/dissertation/dissertation.tex
foo

% Don't let biblatex change any of the casing in the bibliography file
\DeclareRobustCommand{\MakeSentenceCase*}[1]{{#1}}

@@

\section{Methods}

@@

Four popular-hashtags datasets were collected at four different time periods, spanning across one month.
Multiple datasets were collected to ensure that the results weren't dependent on a particular sample of hashtags, and would generalize across different popular-hashtag datasets and different time periods.
Each dataset required around four days to collect enough tweets to reach the quota, and one collection was done each week.

@@

Multiple samples were taken across a broad range of user types for each dataset to ensure that the model results weren't dependent on specific attributes in a single sample.
For the StackOverflow popular-users dataset, a total of 12 slices of users were taken at different levels of reputation and total number of questions asked.
500 users were sampled that had a reputation higher than \num{500}, \num{1000}, \num{5000}, \num{10000}, \num{50000}, and \num{100000}.
100 users were sampled that had asked a total number of questions higher than \num{50}, \num{100}, \num{200}, \num{300}, \num{400}, \num{500}.

@@

So it may be worthwhile to use the standard form instead of the optimized learning form when declarative memory is modeled across a broader range of tasks.
However, this shouldn't be taken as a strong recommendation,
since the performance benefits seen here could simply be due to the fact that the declarative memory retrievals span across a large period of time, which isn't the case for all tasks.
Nonetheless, the only real reason not to use the standard form is due to computational resource limitations.
Since this becomes less of an issue as time progresses, it may be worthwhile to reconsider the default value used in ACT-R.
It may be the case that the optimized learning form of the equation is no longer computationally necessary.
And if the standard form is used as the default (or simply tried in another experiment),
the results from this study suggest that the optimal decay rate value should be set slightly higher (0.8) than the default when optimized learning is used (0.5).

@@

\subsection{Methods}

@@

However, this relaxed metric was used for the rest of the experiments that involve combining terms because the optimal weights found by the regression are based on using the relaxed metric.
If these weights are then used to compute the constrained form of model accuracy, biases may be introduced since the regression wasn't optimized to the constrained form, and weights may change if that was done.
Also, there is no current logistic regression technique in R that can be constrained to label a specific number of 1s for each observation. 
To optimize parameters based on the constrained form,
a full-enumeration numerical approach would have to be developed that systematically searches the space of coefficient weights and uses a custom objective fitting function to find the best fit.
However, this was not pursued for three reasons:
Around \num{2500} total regressions were run on these datasets,
a custom implementation in R would not likely run in compiled C,
and even if it was implemented in a compiled language it would be computationally intractable without a closed-form solution.

@@

\section{Results for Combining Terms}

@@

Looking across the plots, including the offset significantly increases model accuracy, for all models, terms, and sites.
Accuracy is improved even when only a single term is used.
Although in this case the weight of the term is irrelevant (only one term, weighting for that term doesn't matter),
using an offset still improves accuracy since the offset term ensures that only a small number of tags per posts are predicted by the model to be the chosen tags.

@@

>>>>

% Do not let biblatex change any of the casing in the bibliography file
\DeclareRobustCommand{\MakeSentenceCase*}[1]{{#1}}

@@

\section{Method}

@@

Four popular-hashtags datasets were collected at four different time periods, spanning across one month.
Multiple datasets were collected to ensure that the results were not dependent on a particular sample of hashtags, and would generalize across different popular-hashtag datasets and different time periods.
Each dataset required around four days to collect enough tweets to reach the quota, and one collection was done each week.

@@

Multiple samples were taken across a broad range of user types for each dataset to ensure that the model results were not dependent on specific attributes in a single sample.
For the StackOverflow popular-users dataset, a total of 12 slices of users were taken at different levels of reputation and total number of questions asked.
500 users were sampled that had a reputation higher than \num{500}, \num{1000}, \num{5000}, \num{10000}, \num{50000}, and \num{100000}.
100 users were sampled that had asked a total number of questions higher than \num{50}, \num{100}, \num{200}, \num{300}, \num{400}, \num{500}.

@@

So it may be worthwhile to use the standard form instead of the optimized learning form when declarative memory is modeled across a broader range of tasks.
However, this should not be taken as a strong recommendation,
since the performance benefits seen here could simply be due to the fact that the declarative memory retrievals span across a large period of time, which is not the case for all tasks.
Nonetheless, the only real reason not to use the standard form is due to computational resource limitations.
Since this becomes less of an issue as time progresses, it may be worthwhile to reconsider the default value used in ACT-R.
It may be the case that the optimized learning form of the equation is no longer computationally necessary.
And if the standard form is used as the default (or simply tried in another experiment),
the results from this study suggest that the optimal decay rate value should be set slightly higher (0.8) than the default when optimized learning is used (0.5).

@@

@@

However, this relaxed metric was used for the rest of the experiments that involve combining terms because the optimal weights found by the regression are based on using the relaxed metric.
If these weights are then used to compute the constrained form of model accuracy, biases may be introduced since the regression was not optimized to the constrained form, and weights may change if that was done.
Also, there is no current logistic regression technique in R that can be constrained to label a specific number of 1s for each observation. 
To optimize parameters based on the constrained form,
a full-enumeration numerical approach would have to be developed that systematically searches the space of coefficient weights and uses a custom objective fitting function to find the best fit.
However, this was not pursued for three reasons:
Around \num{2500} total regressions were run on these datasets,
a custom implementation in R would not likely run in compiled C,
and even if it was implemented in a compiled language it would be computationally intractable without a closed-form solution.

@@

\section{Results for Combining Predictors}

@@

Looking across the plots, including the offset significantly increases model accuracy, for all models, terms, and sites.
Accuracy is improved even when only a single term is used.
Although in this case the weight of the term is irrelevant (only one term, weighting for that term does not matter),
using an offset still improves accuracy since the offset term ensures that only a small number of tags per posts are predicted by the model to be the chosen tags.

@@

The Bayesian and vector-based memory models differ in how the co-occurrence matrix is represented (lossy compression for vector-based models).
They also differ in how the activation of each tag is calculated, given the different representations.
The vector-based memory models, like LSA, compute a correlation between the context and each tag, while the Bayesian models sum the log likelihoods that each word in the context occurs with each tag.
These two methods for computing the activation are correlated but not identical.
Further research could try to tease apart how model accuracy is independently influenced by the representation (compressed or not) and the activation calculation (correlation or log likelihood).
To do this, four models could be constructed (uncompressed and compressed representation crossed with correlation and log likelihood activation) compared to the two in this study,
so that the effects of representation and activation calculation could be partialed out.
This would allow, for example, the method of compression for the random permutation vector-based model to be analyzed independent of activation calculation.
The results would provide additional evidence to help determine the specific representation that the declarative memory system uses to represent co-occurrences,
and also the specific activation calculation that is carried out by the system.
